---
title: "DS311 - R Lab Assignment"
author: "Gabrielle Salamanca"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: united
    highlight: tango
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Assignment 1

* In this assignment, we are going to apply some of the build in data set in R for descriptive statistics analysis.
* To earn full grade in this assignment, students need to complete the coding tasks for each question to get the result.
* After finished all the questions, knit the document into HTML format for submission.

### Question 1

Using the **mtcars** data set in R, please answer the following questions.

```{r}
# Loading the data
data(mtcars)

# Head of the data set
head(mtcars)
```

a. Report the number of variables and observations in the data set.

```{r}
# Enter your code here!
dim(mtcars)
# Answer:
print("There are total of 11 variables and 32 observations in this data set.")
```

b. Print the summary statistics of the data set and report how many discrete and continuous variables are in the data set.

```{r}
# Enter your code here!
summary(mtcars)
# Answer:
print("There are 5 discrete variables and 6 continuous variables in this data set.")
```

c. Calculate the mean, variance, and standard deviation for the variable **mpg** and assign them into variable names m, v, and s. Report the results in the print statement.

```{r}
# Enter your code here!
mpg <- mtcars$mpg
m <- mean(mpg)
v <- var(mpg)
s <- sqrt(v)
# print
print(paste("The average of Mile Per Gallon from this data set is ",  m, " with variance ",  v, " and standard deviation",  s, "."))
```

"The average of Mile Per Gallon from this data set is  20.090625  with variance  36.3241028225806  and standard deviation 6.0269480520891 ."

d. Create two tables to summarize 1) average mpg for each cylinder class and 2) the standard deviation of mpg for each gear class.

```{r}
# Enter your code here!
cyl <- mtcars$cyl
gear <- mtcars$gear
# table
aggregate(mpg ~ cyl, mtcars, mean)
aggregate(mpg ~ gear, mtcars, sd)
```

e. Create a crosstab that shows the number of observations belong to each cylinder and gear class combinations. The table should show how many observations given the car has 4 cylinders with 3 gears, 4 cylinders with 4 gears, etc. Report which combination is recorded in this data set and how many observations for this type of car.

```{r}
# Enter your code here!
table(cyl, gear)
print("The most common car type in this data set is car with 8 cylinders and 3 gears. There are total of 12 cars belong to this specification in the data set.")
```

"The most common car type in this data set is car with 8 cylinders and 3 gears. There are total of 12 cars belong to this specification in the data set."

*** 

### Question 2

Use different visualization tools to summarize the data sets in this question.

a. Using the **PlantGrowth** data set, visualize and compare the weight of the plant in the three separated group. Give labels to the title, x-axis, and y-axis on the graph. Write a paragraph to summarize your findings.

```{r}
# Load the data set
data("PlantGrowth")

# Head of the data set
head(PlantGrowth)

# Enter your code here!
boxplot(weight ~ group, data = PlantGrowth,
        xlab = "Group", ylab = "Weight",
        main = "Weight by Group")
```

Result:

=> Report a paragraph to summarize your findings from the plot!

In this box plot, the median of each group is quite different. We see that the control group has a higher median than the two treated ones, and treatment 1 has a higher median than treatment 2. We also see that the first treatment has more variability in weight compared to the other 2, having a noticeable outlier. So the different fertilizer treatments do have an effect, but depending on what you desire, treatment 2 would be a good fertilizer to use, because the plants have a higher weight despite having a lower median.

b. Using the **mtcars** data set, plot the histogram for the column **mpg** with 10 breaks. Give labels to the title, x-axis, and y-axis on the graph. Report the most observed mpg class from the data set.

```{r}
hist(mpg, breaks = 10, main = "MPG of mtcars", xlab = "MPG", ylab = "Frequency")
print("Most of the cars in this data set are in the class of 15 mile per gallon.")
```

c. Using the **USArrests** data set, create a pairs plot to display the correlations between the variables in the data set. Plot the scatter plot with **Murder** and **Assault**. Give labels to the title, x-axis, and y-axis on the graph. Write a paragraph to summarize your results from both plots.

```{r}
# Load the data set
data("USArrests")

# Head of the data set
head(USArrests)

# Enter your code here!
murder <- USArrests$Murder
assault <- USArrests$Assault
plot(murder, assault, xlab = "Murder", ylab = "Assault", main = "Murder and Assault Arrests")
abline(lm(assault ~ murder), col = "darkorchid")
```

Result:

=> Report a paragraph to summarize your findings from the plot!

As murder arrests increases, so does assault arrests. We see that there is a positive correlation between these two variables. There are also plenty of points that stray from the line. Some of them show that assault is high when murder is low and vice versa, but considering this dataset has 1973 statistics on murder, assault, and rape arrests per 100,000 population in each of the 50 states, there will be variability across each state.

***

### Question 3

Download the housing data set from www.jaredlander.com and find out what explains the housing prices in New York City.

Note: Check your working directory to make sure that you can download the data into the data folder. 

```{r, echo=FALSE}
# Load and clean the housing data set 
download.file(url='https://www.jaredlander.com/data/housing.csv',
              destfile='data/housing.csv', mode='wb')
housingData <- read.csv('data/housing.csv')
housingData <- subset(housingData, 
                      select = c("Neighborhood", "Market.Value.per.SqFt", "Boro", "Year.Built"))
housingData <- na.omit(housingData)                   
```

a. Create your own descriptive statistics and aggregation tables to summarize the data set and find any meaningful results between different variables in the data set.

```{r}
# Head of the cleaned data set
head(housingData)

# Enter your code here!

neighbor <- as.factor(housingData$Neighborhood)
market <- housingData$Market.Value.per.SqFt
boro <- as.factor(housingData$Boro)
year <- housingData$Year.Built
# summary
summary(housingData)
summary(neighbor)
summary(boro)
# table
aggregate(market, by = list(boro), FUN = function(x) c(mean = mean(x), sd = sd(x)))
table(neighbor)
table(neighbor, boro)
table(neighbor, cut(year, breaks = seq(min(year), max(year), by = 10)))
```

b. Create multiple plots to demonstrates the correlations between different variables. Remember to label all axes and give title to each graph.

```{r}
# Enter your code here!

library(ggplot2)
library(dplyr)
# variable
avgMarket <- tapply(market, boro, mean)
# plot
ggplot(data = housingData, aes(x = year, y = market)) +
  geom_point() +
  labs(title = "Market Value per Sqft vs Year Built",
       x = "Year Built",
       y = "Market Value per Sqft")

ggplot(data = housingData, aes(x = boro, y = market)) +
  geom_boxplot() +
  labs(title = "Market Value per Sqft by Borough",
       x = "Borough",
       y = "Market Value per Sqft")

barplot(avgMarket, 
        main = "Average Market Value per Sqft by Borough", 
        xlab = "Borough", 
        ylab = "Average Market Value per Sqft")

```

c. Write a summary about your findings from this exercise.

=> Enter your answer here!

The bar plot and histogram appear quite similar. In both, Manhattan is a popular borough that has the highest market value per square feet, completely dwarfing all the other boroughs. Also, Staten Island has the lowest market value per square feet, barely appearing in the histogram. In the scatter plot shows that most of the market value per square feet is between 50 and 250, especially more densely for those built between 1975 and 2000s. Those built in the 2000s do seem to have a higher reach in the market value compared to older homes, but of course there will be outliers. From 1825 to 1900, there aren't many housings being put up on market, and seem to keep between 100 and 200 market value per square feet.
